A standardized framework for risk-based assessment of treatment effect heterogeneity in observational healthcare databases

### 2021-09-07 submission/manuscript.rmd
align results section to methods for prediction step

  - A single treatment cohort ($T$) which includes patients with disease $D$
    receiving the target treatment of interest. For example, a set of
    hypertension patients within a database that receive angiotensin-converting
    enzyme inhibitors, followed from the time of initiation until the time of
    censoring.
  - A single comparator cohort ($C$) which includes patients with disease $D$
    receiving the comparator (control) treatment. For example, a set of patients
    in a database that receive beta blockers, followed from the time of
    initiation until the time of censoring.
  - One or more outcome cohorts ($O_1,\dots,O_n$) that contain patients
    developing the outcomes of interest. For example, the set of patients in a
    database that have at least one occurrence of acute myocardial infarction in
    their record.

@@

  - Treatment cohort: Patients receiving any drug within the ACE-inhibitor class
    with at least one year of follow-up before treatment initiation and a
    recorded hypertension diagnosis within that year.
  - Comparator cohort: Patients receiving any drug within the beta blocker class
    with at least one year of follow-up before treatment initiation and a
    recorded hypertension diagnosis within that year.
  - Outcome cohorts: We consider 3 main and 6 safety outcome cohorts. These are
    patients in the database with a diagnosis of: acute myocardial infarction
    (MI); hospitalization with heart failure; ischemic or hemorrhagic stroke
    (efficacy outcomes); hypokalemia; hyperkalemia; hypotension; angioedema;
    cough; abnormal weight gain (safety outcomes).

@@

@@

We developed separate prediction models for all efficacy outcomes (acute MI,
hospitalization with heart failure and hemorrhagic or ischemic stroke) in each
database.

@@

The prediction models were estimated on the propensity score matched (1:1) subset of
the population, using caliper of 0.2 and after excluding patients having the
outcomes any time prior to treatment initiation. We chose a time horizon of 2
years after inclusion into the target cohort. For this demonstration, we
developed the prediction models using LASSO logistic regression with 3-fold
cross validation for hyper-parameter selection. We chose a time horizon of 2
years after inclusion into the target cohort. For this demonstration, we
developed the prediction models using LASSO logistic regression with 3-fold
cross validation for hyper-parameter selection.

@@

For each outcome and in each risk stratum there were adequate numbers of
patients (Table \@ref(tab:table)). The discriminative ability of the prediction
models was moderate in the matched development subset (c-index 0.76 for acute MI
; 0.79 for hospitalization with heart failure; 0.74 for stroke; ), in the
general population (c-index 0.74 for acute MI; 0.77 for hospitalization with
heart failure; 0.73 for stroke), in the treatment cohort (c-index for acute MI
it was 0.71, for hospitalization with heart failure was 0.76 and for stroke it
was 0.72) and in the comparator cohort (c-index for acute MI it was 0.79 for
hospitalization with heart failure was 0.79 and for stroke it was 0.75).
\captionsetup{width=\textwidth}
```{r table, echo=FALSE, warning=FALSE, message=FALSE, results='asis', booktabs=TRUE}
table <- readr::read_csv(
  here::here("data/processed/table1.csv"),
  na = character()
) %>%
  select(-database)

knitr::kable(
  table,
  format = "latex",
  longtable = TRUE,
  booktabs = TRUE,
  col.names = c(
    "Outcome", "Risk quarter",
    "Persons", "Person years", "Events",
    "Persons", "Person years", "Events"
  ),
   format.args = list(
    big.mar = ","
  ),
  caption = "Number of patients, person years and events within quarters of predicted risk for hospitalization with heart failure for the 3 main outcomes of the study (acute myocardial infarction, hospitalization with heart failure and ischemic or hemorrhagic stroke) across the 3 databases."
) %>%
  kableExtra::group_rows("CCAE", 1, 12) %>%
  kableExtra::group_rows("MDCD", 13, 24) %>%
  kableExtra::group_rows("MDCR", 25, 36) %>%
  kableExtra::add_header_above(
    c(
      "", "", 
      "ACE inhibitors" = 3,
      "Beta blockers"  = 3
    )
  )
```

```{r tableAucs, echo=FALSE, results='asis', booktabs=TRUE}
table <- tibble(
  population = c(
    "Matched set",
    "Treatment cohort",
    "Comparator cohort",
    "Entire population"
  ),
  ccae = c(
    75.78, 70.65, 79.1, 74.06
  ),
  mdcd = c(
    78.12, 78.29, 80.26, 79.46
  ),
  mdcr = c(
    71.09, 69.03, 68.53, 69.07
  )
)

>>>>

  - A single treatment cohort ($T$) which includes patients with disease $D$
    receiving the target treatment of interest. For example, a set of
    hypertension patients within a database that receive angiotensin-converting
    enzyme inhibitors, followed from the time of initiation until the time of
    censoring.
  - A single comparator cohort ($C$) which includes patients with disease $D$
    receiving the comparator (control) treatment. For example, a set of patients
    in a database that receive beta blockers, followed from the time of
    initiation until the time of censoring.
  - One or more outcome cohorts ($O_1,\dots,O_n$) that contain patients
    developing the outcomes of interest. For example, the set of patients in a
    database that have at least one occurrence of acute myocardial infarction
    (MI) in their record.

@@

  - Treatment cohort: Patients receiving any drug within the ACE-inhibitor class
    with at least one year of follow-up before treatment initiation and a
    recorded hypertension diagnosis within that year.
  - Comparator cohort: Patients receiving any drug within the beta blocker class
    with at least one year of follow-up before treatment initiation and a
    recorded hypertension diagnosis within that year.
  - Outcome cohorts: We consider 3 main and 6 safety outcome cohorts. These are
    patients in the database with a diagnosis of: acute MI; hospitalization with
    heart failure; ischemic or hemorrhagic stroke (efficacy outcomes);
    hypokalemia; hyperkalemia; hypotension; angioedema; cough; abnormal weight
    gain (safety outcomes).

@@

Our analyses included a total of 784,561, 66,820 and 101,661 patients initiating
treatment with ACE inhibitors and 395,740, 45,999 and 69,798 patients initiating
treatment with beta blockers in CCAE, MDCD and MDCD respectively (Table
\@ref(tab:table)). Adequate numbers of patients were included in all strata of
predicted acute MI risk.

\captionsetup{width=\textwidth}
```{r table, echo=FALSE, warning=FALSE, message=FALSE, results='asis', booktabs=TRUE}
table <- readr::read_csv(
  here::here("data/processed/table1.csv"),
  na = character()
) %>%
  select(-database)

knitr::kable(
  table,
  format = "latex",
  longtable = TRUE,
  booktabs = TRUE,
  col.names = c(
    "Outcome", "Risk quarter",
    "Persons", "Person years", "Events",
    "Persons", "Person years", "Events"
  ),
   format.args = list(
    big.mar = ","
  ),
  caption = "Number of patients, person years and events within quarters of predicted risk for acute MI for the 3 main outcomes of the study (acute MI, hospitalization with heart failure and ischemic or hemorrhagic stroke) across the 3 databases. The populations used for the evaluation of hosptialization with heart failure and stroke are subsets of the populations used for the evaluation of actute MI."
) %>%
  kableExtra::group_rows("CCAE", 1, 12) %>%
  kableExtra::group_rows("MDCD", 13, 24) %>%
  kableExtra::group_rows("MDCR", 25, 36) %>%
  kableExtra::add_header_above(
    c(
      "", "", 
      "ACE inhibitors" = 3,
      "Beta blockers"  = 3
    )
  )
```

@@

We internally developed separate prediction models for acute MI in all 3
databases.

@@

The prediction models were estimated on the propensity score matched (1:1)
subset of the population, using caliper of 0.2 and after excluding patients
having the outcomes any time prior to treatment initiation. We chose a time
horizon of 2 years after inclusion into the target cohort. For this
demonstration, we developed the prediction models using LASSO logistic
regression with 3-fold cross validation for hyper-parameter selection. For this
demonstration, we developed the prediction models using LASSO logistic
regression with 3-fold cross validation for hyper-parameter selection.

The models developed in the 3 databases had moderate discriminative performance
(internally validated) with no major issues of overfitting to any cohort except
for the case of CCAE database in which the derived prediction model performed
better in the comparator cohort (Table \@ref(tab:tableAucs)). We also observed
lower performance of the prediction model developed in MDCR compared to the
other 2 databases. Results on the calibration of the prediction models can be
found in the supplement.

\captionsetup{width=\textwidth}
```{r tableAucs, echo=FALSE, results='asis', booktabs=TRUE}
table <- tibble(
  population = c(
    "Matched set",
    "Treatment cohort",
    "Comparator cohort",
    "Entire population"
  ),
  ccae = c(
    75.78, 70.65, 79.1, 74.06
  ),
  mdcd = c(
    78.12, 78.29, 80.26, 79.46
  ),
  mdcr = c(
    71.09, 69.03, 68.53, 69.07
  )
)

@@

For each outcome and in each risk stratum there were adequate numbers of
patients (Table \@ref(tab:table)). The discriminative ability of the prediction
models was moderate in the matched development subset (c-index 0.76 for acute MI
; 0.79 for hospitalization with heart failure; 0.74 for stroke; ), in the
general population (c-index 0.74 for acute MI; 0.77 for hospitalization with
heart failure; 0.73 for stroke), in the treatment cohort (c-index for acute MI
it was 0.71, for hospitalization with heart failure was 0.76 and for stroke it
was 0.72) and in the comparator cohort (c-index for acute MI it was 0.79 for
hospitalization with heart failure was 0.79 and for stroke it was 0.75).
